### 2021-11-02 paper/pldi22/formal.tex
Minor update
More proof

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog$ is defined as
follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \initst \in \initstset{\prog} \wedge
  \initst \trans{\prog}^* \st \}
\]
where $\trans{\prog}^*$ denotes one or more repetition of $\trans{\prog}$, and
$\st \trans{\prog} \st'$ if and only if $\st = (\lab, \_, \_, \_)$ and
$\sem{\getinst{\prog}(\lab)}(\st) = \st'$. Another way to represent the concrete
semantics is to define a transfer function $\transfer: \powerset{\stset}
\rightarrow \powerset{\stset}$:
\[
  \sem{\prog} = \lim_{n \rightarrow \infty}\transfer^n(\initstset{\prog})
\]
where
\begin{equation}\label{eqn:transfer}
  \transfer(S) = \{ \st' \mid \st \in S \wedge \st \trans{\prog} \st' \}
\end{equation}
Now, we define the denotational semantics of expressions, instructions, and
functions.

@@

An abstract state $\aelem \in \aelemset$ is a triple of a dynamic call set, a
static call set, and a flow-sensitive abstract environment.  A dynamic call set
$\adcall \in \adcallset$ is a set of call labels representing function calls
with multiple functions or no static arguments. A static call set $\ascall \in
\ascallset$ is a set of possible static call edges from labels to function, and
it represent a set of \textit{static calling contexts} $\ascallgamma(\ascall)$
used as the scope of program transformations.  To more precisely in the static
calling contexts, we define a flow-sensitive abstract environment $\afsenv \in
\afsenvset$. It is a mapping from labels to abstract environments to represents
the shape of environments in each label. An abstract environment $\aenv \in
\aenvset$ is a mapping from variables to abstract values to describe values
stored in variables. An abstract value $\aval \in \avalset$ is a static value
$\val$, a syntactic view $\atree$, a dynamic value $\top$, or nothing $\bot$.
Now, we define abstract semantics of expressions, instructions, and programs.

@@

\paragraph{Programs} Finally, we define the abstract semantics $\asem{\prog}:
\atreeset \rightarrow \aelemset$ of a program $\prog$ with a given syntactic
view:
\[
  \asem{\prog}(\atree) = \lim_{n \rightarrow \infty}(\atransfer)^n \circ
  \entryaelem(\atree)
\]
It is the least fixed point of the abstract transfer function $\atransfer$ with
an abstract entry state $\entryaelem$ depending on the given syntactic view
$\atree$:
\begin{itemize}
  \item \underline{Abstract Transfer Function}: $\atransfer: \aelemset
    \rightarrow \aelemset$
    \[
      \atransfer(\aelem) = \aelem \join \sem{\adcall} \wedge
      \bigjoin_{\lab \in \labset}{\asem{\getinst{\prog}(\lab)}(\lab, \aelem)}
    \]
    where $\aelem = (\adcall, \ascall, \afsenv)$ and $\sem{\adcall} = (\adcall,
    \ascall, [\lab \mapsto \aenv_\lab \mid \lab \in \adcall \wedge
    \getinst{\prog}(\lab) = \varx = \expr(\cdots) \wedge \aenv_\lab = [\varx
    \mapsto \top]])$.
  \item \underline{Abstract Entry States}: $\entryaelem:
    \atreeset \rightarrow \aelemset$
    \[
      \entryaelem(\atree) = (\varnothing, \varnothing, \afsenv)
    \]
    where \[
      \begin{array}{lcl}
        \atree &=& \ty_k \langle \cdots \rangle\\

@@

        \ty_k.\eval &=& \lambda \kwdef \; \kwrl \varx_1, \cdots, \varx_n \kwrr
        \; \entrylab\\

@@

\begin{theorem}[Soundness of $\asem{\prog}$]\label{thm:sound-prog}
  For a given syntactic view $\atree \in \atreeset$, the static analysis
  $\asem{\prog}(\atree)$ is sound by satisfying the following condition:
  \[
    \rsem{\prog}(\atree) \subseteq \aelemgamma \circ \asem{\prog}(\atree)
  \]
\end{theorem}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\entryaelem$]\label{thm:sound-entryaelem}
  For a given syntactic view $\atree \in \atreeset$, the abstract entry state
  $\entryaelem(\atree)$ is a sound approximation of the entry states
  $\entrystset(\atree)$ by satisfying the following condition:
  \[
    \entrystset(\atree) \subseteq \aelemgamma \circ \entryaelem(\atree) 
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\atransfer$]\label{thm:sound-atransfer}
  The abstract transfer function $\atransfer$ is a sound approximation of the
  transfer function $\transfer$ by satisfying the following condition:
  \[
    \forall \aelem \in \aelemset. \;
    \transfer \circ \aelemgamma(\aelem) \subseteq
    \aelemgamma \circ \atransfer(\aelem)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\inst}$]\label{thm:sound-inst}
  For a given label $\lab \in \labset$ and its instruction $\inst =
  \getinst{\prog}(\lab)$, the abstract semantics $\asem{\inst}$ of the
  instruction is a sound approximation of its concrete semantics $\sem{\inst}$:
  \[
    \forall \aelem \in \aelemset. \;
    \st \in \aelemgamma(\aelem) \wedge \st = (\lab, \_, \_, \_) \Rightarrow
    \sem{\inst}(\st) \in \aelemgamma \circ \asem{\inst}(\lab, \aelem)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\expr}$]\label{thm:sound-inst}
  The abstract semantics $\asem{\inst}$ of an expression $\expr$ is a sound
  approximation of its concrete semantics $\sem{\expr}$:
  \[
    \forall \aenv \in \aenvset. \;
    \st \in \aenvgamma(\aenv) \Rightarrow
    \sem{\expr}(\st) \in \avalgamma \circ \asem{\expr}(\aenv)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

@@

\begin{theorem}[Semantics Preservation of $\transform{\prog}$]
  \label{thm:preserve-prog}
  For a given syntactic view $\atree \in \atreeset$, a transformed program
  $\prog' = \transform{\prog}(\aelem)$ with an abstract state $\aelem =
  (\adcall, \ascall, \afsenv) \in \aelemset$ preserves the restricted semantics
  $\rsem{\prog}(\atree)$ under the static calling contexts
  $\ascallgamma(\ascall)$ if $\aelem$ is sound:
  \[
    \todo
    % \rsem{\prog}(\atree) \subseteq \aelemgamma(\aelem) \Rightarrow
    % \forall \st \in \rsem{\prog}(\atree) \cap \ascallgamma(\ascall). \;
    % (\st \trans{\prog} \st' \wedge \st \trans{\prog'} \st'') \Rightarrow (\st' =
    % \st'')
  \]
\end{theorem}
\begin{proof}
  \todo
\end{proof}

>>>>

The concrete semantics $\sem{\prog}$ of an $\ires$ program $\prog$ is defined as
follows:
\[
  \sem{\prog} = \{ \st \in \stset \mid \exists \initst \in \initstset{\prog}. \;
  \initst \trans{\prog}^* \st \}
\]
where $\trans{\prog}^*$ denotes one or more repetition of $\trans{\prog}$, and
$\st \trans{\prog} \st'$ if and only if $\st = (\lab, \_, \_, \_)$ and
$\sem{\getinst{\prog}(\lab)}(\st) = \st'$. Another way to represent the concrete
semantics is to define a transfer function $\transfer: \powerset{\stset}
\rightarrow \powerset{\stset}$:
\[
  \sem{\prog} = \lim_{n \rightarrow \infty}\transfer^n(\initstset{\prog})
\]
where
\begin{equation}\label{eqn:transfer}
  \transfer(S) = \{ \st' \in \stset \mid \exists \st \in S. \; \st \trans{\prog}
  \st' \}
\end{equation}
Now, we define the denotational semantics of expressions, instructions, and
functions.

@@

An abstract state $\aelem \in \aelemset$ is a triple of a dynamic call set, a
static call set, and a flow-sensitive abstract environment.  A dynamic call set
$\adcall \in \adcallset$ is a set of call labels representing function calls
with multiple functions or no static arguments. A static call set $\ascall \in
\ascallset$ is a set of possible static call edges from labels to function, and
it represent a set of \textit{static calling contexts}
$\ascallgamma(\ascall)(\atree)$ with a given syntactic view $\atree$ to
determine the scope of program transformations.  To more precisely in the static
calling contexts, we define a flow-sensitive abstract environment $\afsenv \in
\afsenvset$. It is a mapping from labels to abstract environments to represents
the shape of environments in each label. An abstract environment $\aenv \in
\aenvset$ is a mapping from variables to abstract values to describe values
stored in variables. An abstract value $\aval \in \avalset$ is a static value
$\val$, a syntactic view $\atree$, a dynamic value $\top$, or nothing $\bot$.
Now, we define abstract semantics of expressions, instructions, and programs.

@@

\paragraph{Programs} Finally, we define the abstract semantics $\asem{\prog}:
\atreeset \rightarrow \aelemset$ of a program $\prog$ with a given syntactic
view:
\[
  \asem{\prog}(\atree) = \lim_{n \rightarrow \infty}(\atransfer)^n \circ
  \entryaelem(\atree)
\]
It is the least fixed point of the abstract transfer function $\atransfer$ with
an abstract entry state $\entryaelem$ depending on the given syntactic view
$\atree$:
\begin{itemize}
  \item \underline{Abstract Transfer Function}: $\atransfer: \aelemset
    \rightarrow \aelemset$
    \[
      \atransfer(\aelem) = \aelem \join \sem{\adcall} \join
      \bigjoin_{\lab \in \labset}{\asem{\getinst{\prog}(\lab)}(\lab, \aelem)}
    \]
    where $\aelem = (\adcall, \ascall, \afsenv)$ and $\sem{\adcall} = (\adcall,
    \ascall, [\lab \mapsto \aenv_\lab \mid \lab \in \adcall \wedge
    \getinst{\prog}(\lab) = \varx = \expr(\cdots) \wedge \aenv_\lab = [\varx
    \mapsto \top]])$.
  \item \underline{Abstract Entry States}: $\entryaelem:
    \atreeset \rightarrow \aelemset$
    \[
      \entryaelem(\atree) = (\varnothing, \varnothing, \afsenv)
    \]
    where \[
      \begin{array}{lcl}
        \atree &=& \ty_k \langle \cdots \rangle\\

@@

        \ty_k.\eval &=& \kwdef \; \kwrl \varx_1, \cdots, \varx_n \kwrr
        \; \entrylab\\

@@

\begin{theorem}[Soundness of $\asem{\prog}$]\label{thm:sound-prog}
  For a given syntactic view $\atree \in \atreeset$, the static analysis
  $\asem{\prog}(\atree)$ is sound by satisfying the following condition:
  \[
    \rsem{\prog}(\atree) \subseteq \aelemgamma \circ \asem{\prog}(\atree)
  \]
\end{theorem}
\begin{proof}
  We use an induction on $n$.  Lemma~\ref{lem:sound-entryaelem} proves the base
  case, $n=0$. Then, we can prove the inductive case using
  Lemma~\ref{lem:sound-atransfer}.
\end{proof}

\begin{lemma}[Soundness of $\entryaelem$]\label{lem:sound-entryaelem}
  For a given syntactic view $\atree \in \atreeset$, the abstract entry state
  $\entryaelem(\atree)$ is a sound approximation of the entry states
  $\entrystset(\atree)$ by satisfying the following condition:
  \[
    \entrystset(\atree) \subseteq \aelemgamma \circ \entryaelem(\atree)
  \]
\end{lemma}
\begin{proof}
  Since all the conditions described in the definition $\entrystset$ and
  $\entryaelem$ are exactly same with a given syntactic view $\atree$, this
  lemma is trivially proved.
\end{proof}

\begin{lemma}[Soundness of $\atransfer$]\label{lem:sound-atransfer}
  The abstract transfer function $\atransfer$ is a sound approximation of the
  transfer function $\transfer$ by satisfying the following condition:
  \[
    \forall \aelem \in \aelemset. \;
    \transfer \circ \aelemgamma(\aelem) \subseteq
    \aelemgamma \circ \atransfer(\aelem)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\inst}$]\label{lem:sound-inst}
  \todo
  % For a given label $\lab \in \labset$ and its instruction $\inst =
  % \getinst{\prog}(\lab)$, the abstract semantics $\asem{\inst}$ of the
  % instruction is a sound approximation of its concrete semantics $\sem{\inst}$:
  % \[
  %   \forall \aelem = \in \aelemset. \;
  %   \st \in \aelemgamma(\aelem) \wedge \st = (\lab, \_, \_, \_) \Rightarrow
  %   \sem{\inst}(\st) \in \aelemgamma \circ \asem{\inst}(\lab, \aelem)
  % \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

\begin{lemma}[Soundness of $\asem{\expr}$]\label{lem:sound-inst}
  The abstract semantics $\asem{\inst}$ of an expression $\expr$ is a sound
  approximation of its concrete semantics $\sem{\expr}$:
  \[
    \forall \aenv \in \aenvset. \;
    \st \in \aenvgamma(\aenv) \Rightarrow
    \sem{\expr}(\st) \in \avalgamma \circ \asem{\expr}(\aenv)
  \]
\end{lemma}
\begin{proof}
  \todo
\end{proof}

@@

\begin{theorem}[Semantics Preservation of $\transform{\prog}$]
  \label{thm:preserve-prog}
  For a given syntactic view $\atree \in \atreeset$, a transformed program
  $\prog' = \transform{\prog}(\aelem)$ with an abstract state $\aelem =
  (\adcall, \ascall, \afsenv) \in \aelemset$ preserves the restricted semantics
  $\rsem{\prog}(\atree)$ under the static calling contexts
  $\ascallgamma(\ascall)(\atree)$ if $\aelem$ is sound:
  \[
    \todo
    % \rsem{\prog}(\atree) \subseteq \aelemgamma(\aelem) \Rightarrow
    % \forall \st \in \rsem{\prog}(\atree) \cap \ascallgamma(\ascall). \;
    % (\st \trans{\prog} \st' \wedge \st \trans{\prog'} \st'') \Rightarrow (\st' =
    % \st'')
  \]
\end{theorem}
\begin{proof}
  \todo
\end{proof}
